---
title: "Nordic Exceptionalism"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(patchwork)
library(ggrepel)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

(Add info on idea of "nordic exceptionalism" and outline of anaylsis/exploration)

The Nordic Region includes five countries: Denmark, Norway, Sweden, Finland, and Iceland. 

Generating dataset with nordic countries and overall group averages.

```{r warnings = FALSE}
# Creating nordic only dataset
nordic_only = 
  read_csv("./data/merged_df.csv") %>% 
  filter(countries %in% c("Denmark", "Norway", "Sweden", 
           "Finland", "Iceland")) %>% 
  select(-X1) %>% 
  rename("overall_happiness_score" = "ladder_score")

# Creating dataset with all other countries
global_other = read_csv("./data/merged_df.csv") %>% 
  select(-X1) %>% 
  mutate(
    countries = str_remove(
      countries, "Denmark|Norway|Sweden|Finland|Icedland")
    ) %>% 
  rename("overall_happiness_score" = "ladder_score")

#Getting averages across dataframes
nordic_averages = summarise_all(nordic_only, mean)
global_averages = summarise_all(global_other, mean, na.rm = TRUE)

#Removing NAs in 'countries' from last step 
nordic_averages[is.na(nordic_averages)] = 0
global_averages[is.na(global_averages)] = 1

#Binding dataframes                          
nordic_data = 
  rbind(nordic_only, nordic_averages, global_averages) %>% 
  mutate(
    countries = as.character(countries),
    countries = 
      str_replace(countries, "0", "nordic_average"),
    countries = 
      str_replace(countries, "1", "other_global_average")
    )

```

Nordic countries vs rest of the world on happiness and freedom scores. Note that "human freedom score" is a composite score of personal and economic freedom. 

```{r}
nordic_data %>% 
  select(countries, overall_happiness_score, human_freedom_score, personal_freedom_score, economic_freedom_score) %>% 
  knitr::kable()
```

Nordic countries vs rest of the world with additionally included sub freedoms.

```{r}
nordic_data %>% 
  select(countries, overall_happiness_score, human_freedom_score, religion, association, assembly, security_safety, rule_of_law, homicide, size_of_government) %>%
  knitr::kable()
```

Plan for two sample t-test comparing Nordic happiness and human freedom score versus the rest of the world. Tables showing summary data of means, SD, and n for nordic vs other countries.

```{r}

#Reformatting data for tests 

merged_setup = read_csv("./data/merged_df.csv") %>% 
  select(-X1) %>% 
  rename("overall_happiness_score" = "ladder_score") %>%
  select(countries, overall_happiness_score,
         human_freedom_score) %>% 
  mutate(
    countries = case_when(
      countries == "Denmark"     ~ "nordic",
      countries == "Norway"      ~ "nordic",
      countries == "Sweden"      ~ "nordic",
      countries == "Finland"     ~ "nordic",
      countries == "Iceland"     ~ "nordic"
    )) %>% 
  mutate(countries = as.character(countries))

merged_setup$countries[is.na(merged_setup$countries)] = "other"

#Group means, counts, and SD for happiness score
group_by(merged_setup, countries) %>%
  summarise(
    count = n(),
    mean = mean(overall_happiness_score, na.rm = TRUE),
    sd = sd(overall_happiness_score, na.rm = TRUE)
  )

#Group means, counts, and SD for freedom score
group_by(merged_setup, countries) %>%
  summarise(
    count = n(),
    mean = mean(human_freedom_score, na.rm = TRUE),
    sd = sd(human_freedom_score, na.rm = TRUE)
  )
```

t-test for overall happiness nordic vs other. Assume unequal variance due to difference in sample size. 

```{r}
t.test(overall_happiness_score ~ countries, 
              data = merged_setup, var.equal = FALSE) %>% 
  broom::tidy() %>% 
  select(estimate:estimate2, statistic, p.value) %>% 
  knitr::kable()
```

t-test for human freedom score nordic vs other. Assume unequal variance due to difference in sample size. 

```{r}
t.test(human_freedom_score ~ countries, 
              data = merged_setup, var.equal = FALSE) %>% 
  broom::tidy() %>% 
  select(estimate:estimate2, statistic, p.value) %>% 
  knitr::kable()
```

To-do: potentially scatter plots with sub freedoms nordic countries 
