---
title: "Nordic Exceptionalism"
output: github_document
---

Change output later 

```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(plotly)

library(ggplot2)
library(grid)
library(rworldmap)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

The Nordic Region includes five countries: Denmark, Norway, Sweden, Finland, and Iceland. Tables comparing the Nordic country, as well as the averages from the Nordic region and global averages for each score. Note that the human freedom score is a composite score of the personal and economic freedom scores. 

```{r warnings = FALSE}
nordic_only = 
  read_csv("./data/merged_df.csv") %>% 
  filter(countries %in% c("Denmark", "Norway", "Sweden", 
           "Finland", "Iceland")) %>% 
  select(-X1)
```


```{r warnings = FLASE}
merged_df = read_csv("./data/merged_df.csv") %>% 
  select(-X1)

nordic_averages = summarise_all(nordic_only, mean)
global_averages = summarise_all(merged_df, mean, na.rm = TRUE)

nordic_averages[is.na(nordic_averages)] = 0
global_averages[is.na(global_averages)] = 1
                          
nordic_data = 
  rbind(nordic_df, nordic_averages, global_averages) %>% 
  mutate(
    countries = as.character(countries),
    countries = 
      str_replace(countries, "0", "nordic_average"),
    countries = 
      str_replace(countries, "1", "global_average")
    )

```

```{r}
nordic_data %>% 
  select(countries, ladder_score, human_freedom_score, personal_freedom_score, economic_freedom_score) %>% 
  knitr::kable()

nordic_data %>% 
  select(countries, ladder_score, human_freedom_score, religion, association, assembly, security_safety, rule_of_law, homicide, size_of_government) %>% 
  knitr::kable()
```


plan - t/test with nordic vs 'rest of world' on happiness and freedom score
plots scatter plots with sub freedoms 
