---
title: "Report"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
# Final Report

This page contains the written report for the final project by Hannah Briggs, Megan Tebbenhoff, and Henry Stoddard.

## Motivation
 In the year 2020, happiness seems to be at an all-time low. A global pandemic rages on into the new year, with cases and deaths surging to levels higher than ever seen before. Such a situation has forced many people in society to give up freedoms and individual liberties previously guaranteed in everyday life. Lockdowns, mask mandates, reduced capacity in stores and hybrid learning have all been imposed to keep people safe and healthy, yet some people see this as an infringement on their personal rights. This connection between freedom and happiness motivated our group to investigate further. 
 While we could not obtain data from 2020, we were able to obtain the most recent data possible (from 2019) surrounding global measures of happiness and freedom. Our goals were to further investigate the relationship between happiness and human freedom, with special attention towards those countries that had the actual highest and lowest scores, and countries that are colloquially thought of as the happiest and most free countries.
 
## Related Work
This exploration of two different metrics came from Megan's experience in her undergraduate courses in international relations. Jeffrey Sachs, a famous economist and policy analyst who wrote a lot about globalization and other international relations topics, is one of the lead minds on the World Happiness Report. The Human Freedom Index is a new measurement tool to describe the relative freedoms of individuals, both economic and personal, living in any given country in the world. Megan thought it could be interesting to look at happiness scores as they relate to human freedom scores, combining the two datasets that she became familiar with in other classes.

## Initial Questions
Our primary questions were:  

* How does happiness vary across nations?  

* How does freedom vary across nations?  

* Does freedom have an impact on happiness/do freedom and happiness have a positive association?

These questions evolved over the course of the project after we did initial mappings and exploratory analysis. We felt it would be interesting to take a closer look at the highest scoring and lowest scoring countries for each of the two metrics to discover any similarities or differences that may be present. We also considered the idea of "Nordic exceptionalism" and wanted to investigate how these measures are similar or different within the Nordic region. In addition, we wanted to see how different subtypes of freedom were connected to happiness, not just overall freedom.

## Data

The data for this project comes from the 2019 World Happiness Report and Human Freedom Index, and can be accessed at the following links: 

* [The 2019 World Happiness Report ](https://worldhappiness.report/ed/2019/)

* [The Human Freedom Index - 2019](https://www.cato.org/human-freedom-index-new)

After importing these datasets, we performed some basic tidying (e.g., renaming variables) and selected a majority of the variables in each dataset in order to get a sense of which ones we wanted to include in our analyses. In addition, we ensured that each dataset was arranged by country alphabetically in order to successfully merge the two. We merged using 'full_join' so that all country data was represented in this final dataset. For the purpose of global maps, we wanted to include all countries available in each dataset for visualization. However, in other analyses where we planned to assess the relationship between happiness scores and human freedom scores, countries that did not have data for both indices were to be removed. Finally, we exported this final, merged dataset as a csv to use as a base for all subsequent analysis. 

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)

# Importing and cleaning datasets

whr_df = 
  read_excel("./data/WHR20_Data.xls") %>% 
  janitor::clean_names() %>% 
  arrange(country_name) %>% 
  rename(countries = country_name) %>% 
  select(-regional_indicator, -standard_error_of_ladder_score, -upperwhisker, -lowerwhisker)

hfi_df = 
  read_excel("./data/human-freedom-index-2019.xlsx",
             skip = 4, na = "-") %>% 
  janitor::clean_names() %>% 
  filter(year == "2017") %>% 
  arrange(countries) %>% 
  select(countries, human_freedom_score, procedural_justice:personal_freedom_score, government_consumption:economic_freedom_score)

# Merging datasets
merged_df = 
  full_join(whr_df, hfi_df, by = "countries") %>% 
  select(countries, ladder_score, human_freedom_score, personal_freedom_score, economic_freedom_score, 
         religion, association, assembly, movement, inheritance_rights,
         disapperances_conflicts_and_terrorism, rule_of_law, homicide, size_of_government) %>% 
  rename("happiness_score" = "ladder_score")

#Exporting merged dataset to use for our analyses
write.csv(merged_df, "./data/merged_df.csv")

```

## Exploratory Analysis

### World Maps
This analysis can be found at the following link: [World Maps](maps_data_clean.html). 

We first wanted to create world maps to get a visual representation of the distribution of happiness and human freedom scores across the world. Using the merged data, we created two maps, one for each metric. We wanted the maps to be interactive using plotly so viewers could see the exact scores for each country.

This became a challenging undertaking to plot our data on top of a world map, as the easiest way to get a projection of the world map through plotly requires the use of country ISO codes. In order to add these codes to our data, we had to merge in a master dataset from plotly that contained ISO codes and country names. This required us to read in the master dataset and mutate the variable name "country" to match our variable name "countries". We also selected just the countries and the ISO codes from the plotly master dataset, as we are not interested in any of the rest of the variables. Below is the code we used to do this.

```{r, message=FALSE, warning=FALSE}
merged_df = 
  read_csv("./data/merged_df.csv") %>% 
  select(-X1)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv") %>% 
  janitor::clean_names() %>% 
  mutate(countries = country) %>% 
  select(countries, code)
```

Upon loading in the dataset, we noticed that some of the country names in this plotly master dataset were different from the country names in our merged dataset, so we had to find the countries in our dataset that did not exactly match the country name in the plotly dataset and recode them to match the plotly dataset before we can merge the two together. Below is the code we used to do this, including notes on the terms we used in the `str_detect()` step.

```{r, message=FALSE, warning=FALSE, results = 'hide'}
recode_df = anti_join(merged_df, df)

df %>% 
  filter(str_detect(countries, "Congo|Gambia|Ivo|Burma|Cyprus|Palestine|Korea|Trinidad|United|Bahamas|Brunei|Verde|Kyrgy"))

#We used the search term "Congo" to find the labels for the Republic of Congo and the Democratic Republic of the Congo.

#Since Ivory Coast can also be referred to as Cote D'Ivoire, we searched for "Ivo" to try to yield either result.

#Upon first search for Myanmar, no results were found. However, Myanmar is sometimes referred to as Burma, so I searched for Burma as well, which gave us a result.

#Upon searching, there was no label for Palestine in the plotly master dataset.

#We used the search term "United" to find the labels for the UK and the USA.

#Since Cape Verde is also known as Cabo Verde, we searched only for Verde.

#We were not sure why there was a missing value for Kyrgyzstan at first. We realized that some sources keep the s while others do not, so we searched only for "Kyrgy".

#Now we need to overwrite our merged dataset with the recoded country names to ensure that the merge with the plotly dataset will work properly.

merged_df = merged_df %>% 
  mutate(countries = recode(countries, "Democratic Republic of the Congo" = "Congo, Democratic Republic of the", 
  "Gambia" = "Gambia, The", "Ivory Coast" = "Cote d'Ivoire", "Myanmar" = "Burma", "North Cyprus" = "Cyprus", 
  "Republic of the Congo" = "Congo, Republic of the", "South Korea" = "Korea, South", 
  "Trinidad" = "Trinidad and Tobago", "UK" = "United Kingdom", "USA" = "United States", "Bahamas" = "Bahamas, The",
  "Brunei Darussalam" = "Brunei", "Cape Verde" = "Cabo Verde", "Kyrgyztan" = "Kyrgyzstan") 
         )

#We are doing a left_join to keep only the countries in our merged dataset rather than having a bunch of NAs in the map creation step.

map_df = left_join(merged_df, df)

```

This new dataframe includes all of our merged data plus the ISO codes, which then made it very easy to plot the world maps with our data overlayed. 

The world maps provide the interactive visualization of our data that we were hoping to show, allowing viewers to explore the scores of over 150 countries for each metric. We hope that viewers will notice that, while many European and North American countries have higher scores in both metrics, there is a lot more nuance to the scores that will be further explored in our other analyses.

### Top 5 Bottom 5
This analysis can be found at the following link: [Top 5 Bottom 5](top_5_bottom_5.html). 

As we began our exploration of the data, we wondered how countries with similar overall human freedom scores related in the subcategories of human freedom scores. We also wondered if the countries with the highest and lowest scores for each metric would be the same. We decided to explore this by looking at the 5 countries with the highest and lowest scores for human freedom and happiness to see where they were similar and where they were different.

We made individual data frames for each of the four groups (bottom/top happiness/human freedom) with four variables: human freedom score, happiness score, economic freedom score, and personal freedom score. 

For the bottom 5 data frames, the data was sorted so that the lowest score would appear first, while the top 5 score data frames were sorted so the highest scores would appear first. The means and standard deviations are provided in the table as a reference point for the viewer to understand the averages of the four scores and how spread out the data is across the countries.

The bar graphs are included to visually show the variation in freedom scores among the top 5 and bottom 5 countries. We wanted to show that, while these are the 5 countries with the lowest or highest scores, there are still differences between them.

We included separate tables with the mean and standard deviation for the top 5 and bottom 5 countries across both metrics to be able to look at how spread out or close together the countries' values are. This allows us to get a better sense of how similar or how different the countries in the top 5 and bottom 5 groups are from one another.

We also included a scatterplot to visually drive home the differences between the groups. This was a great way to show that, while the scores may be very similar on the metric for which we selected (human freedom score or happiness score), there could be variety in the scores on the other metric. Interestingly, we found a lot more variety among the bottom 5 countries than we did among the top 5 countries.

## Additional Analysis

### Linear analysis
After conducting exploratory analyses, we decided to conduct a linear analysis. This analysis can be found in further detail at the following link: [Linear Analysis](Linear-Analysis.html). 

The central purpose behind this analysis was to observe the relationship between happiness (assessed via ladder score in the world happiness report) and freedom (assessed by multiple measures in the human freedom index). 

To begin, we created a scatterplot, which revealed a mostly positive and linear relationship between happiness and overall freedom. In addition, we created box plots and density plots to look for outliers and normality in ladder score data and overall freedom scores. These plots revealed one outlier for freedom and mostly normal distributions (though less so for freedom). The final step before creating linear models was to assess correlation between happiness and overall freedom, in which we found a strong correlation (r = 0.69).

Prior to creating our models, we planned to include several covariates (like GDP, for example) to control for potential confounding. However, upon digging into the data, we discovered that both happiness and freedom scores actually already accounted for such covariates in their scoring system already. Therefore, we decided to make our linear models only containing variables already in the dataset. 

Our first linear model only included happiness as a function of overall freedom. Results showed a strong, positive relationship with a significant p-value, and normally distributed residuals. This result further supported the conclusion that happiness and freedom are positively associated. However, this model did not show a great degree of granularity, so we decided to create two additional, more specific models.

The second linear model included happiness as a function of economic and personal freedom. This model also revealed a positive relationship between both types of freedom and happiness, though their beta terms were smaller than the overall freedom measure's beta term. Residuals were normally distributed. Interestingly, personal freedom had a larger beta term than economic freedom, suggesting that personal freedoms (like the freedom to move, associate, practice religion, etc.) have a greater impact on happiness than economic freedoms (take that capitalism!).

The final model further broke down freedom into freedom of/from assembly, religion, association, movement, security/safety, inheritance rights, disappearances, conflict, terrorism, rule of law, and homicide. This model gave the greatest degree of granularity, but at the cost of smaller effect sizes. In this model, the only significant terms were rule of law (moderate, positive association) and homicide (small, negative association). However, the disappearances, conflicts, and terrorism term was close to being significant as well. The residuals for these models looked less normal than previous models, raising concerns about the validity of this final model.

### Nordic Exceptionalism
This analysis can be found at the following link: [Nordic Exceptionalism](nordic_except.html). Nordic countries tend to have among the happiest and most satisfied citizens. The phrase “Nordic Exceptionalism” is used to describe this phenomenon and ask what exactly it is about these nations that leads to these results. Such countries are noted to offer high levels of social support to their citizens and display lower measures of inequality - while noting that citizens pay high levels of taxes in order to support the public sector (1). We decided to examine the five Nordic countries (Denmark, Norway, Sweden, Finland, and Iceland), looking at their happiness scores from the 2019 WHR, as well as how measures of overall and sub-indicators of freedom may play into this narrative. 

In addition to looking at human freedom scores (overall score) and measures of personal and economic freedom – we pulled several sub-indicators of personal and economic freedom that go into this score. Specifically, measures of freedom related to religion, association, assembly, rule of law, homicide, disappearances/conflicts/terrorism, inheritance rights, and size of government.  

Based on tabular and graphical descriptive analyses, Nordic countries tended to appear fairly similar to one another on measures of happiness and the various measures of freedom we used. We note that, on average, the ‘personal freedom’ scores for Nordic countries tend to be particularly higher compared to the rest of the world. While these countries also have on average higher ‘economic freedom’ scores, the difference is less pronounced comparatively. Nordic countries also tend to be fairly homogenous when looking at these sub-indicators of freedom. Most pronounced in the data, Nordic countries tended to have on average, lower homicide scores, higher size of government scores, and rule of law scores – compared to the rest of the world. Meaning – less intentional homicides, higher degrees of government consumption and taxes, as well as higher adherence to procedural, civil, and criminal justice. 

As we lack information in this dataset on any within country variance, we were unable to carry out statistical testing to confirm whether or not any one of the Nordic countries may statistically differ to a significant degree from the rest. After visualizing the data, we were comfortable treating the Nordic region as one group based on the observed higher degree of similarity among the five countries. We carried out two sample t-tests comparing Nordic countries to all other countries in the world to see whether mean happiness and mean human freedom scores significantly differed between these groups. After checking for homogeneity of variance in the two groups, we ran these t-tests accounting for unequal variance. Both t-tests yielded p-values of < 0.05, indicating sufficient evidence to say that mean citizen happiness and human freedom between these two groups significantly differ. 

## Discussion
Exploratory analysis revealed a wide variation in overall happiness and freedom scores across the globe. Additionally, it revealed that the top scoring countries on each metric were quite similar to one another on the other metric, while there was considerably more variation among the lowest scoring countries. Hong Kong was an outlier among the highest scoring countries for freedom scores, as they had quite a low happiness score relative to the rest of their group. For the lowest scoring countries on the human freedom index, the countries differed greatly in happiness scores. For the lowest scoring countries on happiness score, there was some variation between the countries on human freedom score, but more interestingly, the average human freedom score was quite high at 6.3 out of a possible 10 and the global average sitting around 6.9.

Linear regression analysis revealed a strong relationship between happiness and overall freedom, especially personal freedom. This makes sense, given that human beings have fought for freedom many times over the course of history. As Patrick Henry once said: "Give me liberty, or give me death". It would seem that greater freedom is associated with greater happiness, and (potentially even more so) less freedom is associated with less happiness.

Nordic countries appeared to be particularly higher on measures of 'personal freedom' to the rest of the world. Considering that Nordic countries are known to implore higher taxes on their citizens to fund the public sector, this would be in line with what we expected to see in this data. Statistical analysis demonstrated that mean citizen happiness in Nordic countries significantly differ from mean citizen happiness in the rest of the world; and that mean human freedom in Nordic countries significantly differs from the rest of the world. 

This study had a handful of limitations. One notable limitation was the use of only one year's worth of data. This made the study essentially cross-sectional, and prevented any assessment of temporal changes. Future studies could build on this one by conducting similar analyses but with multiple years of data and looking at changes over time. In addition, we would consider examining Nordic countries in reference to several different regions of the world (as opposed to an "all other" global average) in future analysis. As well as consider other potential variables to control for in investigating how measures of human freedom may be associated with these countries' high happiness and satisfaction scores.  

References: 

1. Frelle-Petersen, C., Hein, A., & Christiansen, M. (2020). The Nordic social welfare model: Lessons for reform [PDF].
*Deloitte Insights*. https://www2.deloitte.com/content/dam/insights/us/articles/43149-the-nordic-social-welfare-model/DI_The-Nordic-social-welfare-model.pdf