---
title: "Top 5 vs. Bottom 5 Countries"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

As a part of our exploratory analyses, we wanted to examine the relationship between human freedom scores and happiness scores for the highest scoring and lowest scoring countries for each metric. We also wanted to look at the subcomponents of the human freedom scores to see if the top scoring or lowest scoring countries had similarities across these metrics. Below, you will find the tabs for each metric with analysis of the 5 highest and 5 lowest scoring countries for that given metric.

```{r setup, include = F}
library(tidyverse)
library(readxl)
library(patchwork)
library(ggrepel)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r, warning = F, message = F, include = F}
top5_bot5_df = 
  read_csv("./data/merged_df.csv") %>%
  drop_na(happiness_score, human_freedom_score) %>% 
  select(countries:economic_freedom_score)
```

```{r, include = F}
bot5_whr = 
  top5_bot5_df %>% 
  arrange(happiness_score) %>% 
  top_n(happiness_score, n = -5)

bot5_whr_avg = summarise_all(bot5_whr, mean)
bot5_whr_sd = data.frame(rbind(apply(bot5_whr, 2, sd)))

bot5_whr_avg[is.na(bot5_whr_avg)] = 0
bot5_whr_sd[is.na(bot5_whr_sd)] = 1

bot5_whr_merge = 
  rbind(bot5_whr, bot5_whr_avg, bot5_whr_sd) %>% 
  mutate(countries = as.character(countries),
    countries =  str_replace(countries, "0", "bot5_whr_average"),
    countries = str_replace(countries, "1", "bot_whr_sd")
      )
```

```{r, include = F}
top5_whr = 
  top5_bot5_df %>% 
  arrange(happiness_score) %>% 
  top_n(happiness_score, n = 5) %>% 
  arrange(desc(happiness_score))

top5_whr_avg = summarise_all(top5_whr, mean)
top5_whr_sd = data.frame(rbind(apply(top5_whr, 2, sd)))

top5_whr_avg[is.na(top5_whr_avg)] = 0
top5_whr_sd[is.na(top5_whr_sd)] = 1

top5_whr_merge = 
  rbind(top5_whr, top5_whr_avg, top5_whr_sd) %>% 
  mutate(countries = as.character(countries),
    countries =  str_replace(countries, "0", "top5_whr_average"),
    countries = str_replace(countries, "1", "top5_whr_sd")
      )

```

```{r, include = F}
top5_hfi = 
  top5_bot5_df %>% 
  select(countries, human_freedom_score, everything()) %>% 
  arrange(human_freedom_score) %>% 
  top_n(human_freedom_score, n = 5) %>% 
  arrange(desc(human_freedom_score)) 

top5_hfi_avg = summarise_all(top5_hfi, mean)
top5_hfi_sd = data.frame(rbind(apply(top5_hfi, 2, sd)))

top5_hfi_avg[is.na(top5_hfi_avg)] = 0
top5_hfi_sd[is.na(top5_hfi_sd)] = 1

top5_hfi_merge = 
  rbind(top5_hfi, top5_hfi_avg, top5_hfi_sd) %>% 
  mutate(countries = as.character(countries),
    countries =  str_replace(countries, "0", "top5_hfi_average"),
    countries = str_replace(countries, "1", "top5_hfi_sd")
      )
```

```{r, include = F}
bot5_hfi = 
  top5_bot5_df %>% 
  select(countries, human_freedom_score, everything()) %>%
  arrange(human_freedom_score) %>% 
  top_n(human_freedom_score, n = -5) 

bot5_hfi_avg = summarise_all(bot5_hfi, mean)
bot5_hfi_sd = data.frame(rbind(apply(bot5_hfi, 2, sd)))

bot5_hfi_avg[is.na(bot5_hfi_avg)] = 0
bot5_hfi_sd[is.na(bot5_hfi_sd)] = 1

bot5_hfi_merge = 
  rbind(bot5_hfi, bot5_hfi_avg, bot5_hfi_sd) %>% 
  mutate(countries = as.character(countries),
    countries =  str_replace(countries, "0", "bot5_hfi_average"),
    countries = str_replace(countries, "1", "bot5_hfi_sd")
      )
```

## Exploring Top 5 and Bottom 5 Country Scores {.tabset}


### Human Freedom Index Scores

#### Bottom 5 Countries for Human Freedom Scores
```{r, echo = F}
bot5_hfi_merge %>% knitr::kable(digits = 3)
```

As we can see here, the countries with the five lowest human freedom scores have some variety in their happiness scores and in their freedom subscores. For example, Venezuela has the lowest overall human freedom score among the bottom 5 countries, but its happiness score is one of the highest in the group. Interestingly, Venezuela has the highest personal freedom score of the bottom 5 countries, but it has the lowest economic freedom score, keeping its overall score low.

```{r}
bot5_hfi %>% 
select(-happiness_score) %>% 
  pivot_longer(
    human_freedom_score:economic_freedom_score,
    names_to = "freedom_subgroup",
    values_to = "score"
  ) %>% 
  mutate(freedom_subgroup = fct_relevel(freedom_subgroup, "human_freedom_score", "personal_freedom_score", 
                     "economic_freedom_score")) %>% 
  ggplot(aes(x = countries, y = score, fill = freedom_subgroup)) + 
  geom_bar(stat = "identity", color = "black", position = position_dodge()) +
  ylim(min = 0, max = 10) + 
  ggtitle("Comparing Bottom 5 Countries on Freedom Subtypes") +
  xlab("Countries")+
  ylab("Freedom Score")+
  labs(fill = "Freedom Subgroup")
```

As seen above, there is a good amount of variety in freedom subgroups between the bottom 5 countries.

#### Top 5 Countries for Human Freedom Scores
```{r, echo = F}
top5_hfi_merge %>% knitr::kable(digits = 3)
```

There appears to be little variety across these scores for the top 5 countries in the freedom scores. However, Hong Kong is a clear outlier in terms of happiness score.

```{r, warning = F}
top5_hfi %>% 
select(-happiness_score) %>% 
  pivot_longer(
    human_freedom_score:economic_freedom_score,
    names_to = "freedom_subgroup",
    values_to = "score"
  ) %>% 
  mutate(freedom_subgroup = fct_relevel(freedom_subgroup, "human_freedom_score", "personal_freedom_score", 
                     "economic_freedom_score")) %>% 
  ggplot(aes(x = countries, y = score, fill = freedom_subgroup)) + 
  geom_bar(stat = "identity", color = "black", position = position_dodge()) +
  ylim(min = 0, max = 10) + 
  ggtitle("Comparing Top 5 Countries on Freedom Subtypes") +
  xlab("Countries")+
  ylab("Freedom Score")+
  labs(fill = "Freedom Subgroup")
```

The countries look more or less the same in terms of freedom subtypes.

```{r, warning = F, echo = F}
hfi_table = 
  rbind(top5_hfi_merge, bot5_hfi_merge) %>% 
  filter(countries %in% c("top5_hfi_average", "top5_hfi_sd", "bot5_hfi_average", "bot5_hfi_sd")) %>% 
  select(countries, human_freedom_score, happiness_score) %>% 
  pivot_longer(human_freedom_score:happiness_score,
    names_to = "score",
    values_to = "aaah"
  ) %>% 
  pivot_wider(
    names_from = "countries",
    values_from = "aaah"
  )

hfi_table %>% knitr::kable()
```

This table is showing the mean and the standard deviations among the bottom 5 and top 5 countries across the two main metrics. The means for human freedom score among the top 5 and bottom 5 countries are dramatically different, while the means for happiness score are a bit closer to one another and the standard deviations are quite high. The high standard deviations for the happiness scores seem to indicate that the countries' scores for happiness score are more spread out in both the top 5 and bottom 5 groups.

```{r, message = F, warning = F}
hfi_combo = full_join(bot5_hfi, top5_hfi)

hfi_combo %>% 
  mutate(text_label = str_c(countries, "\nFreedom Score:", human_freedom_score, "\nHappiness Score: ", happiness_score)) %>%
  plot_ly(x = ~ human_freedom_score, y = ~happiness_score, type = "scatter", mode = "markers", size = 3, alpha = 0.5, color = ~countries, text = ~text_label, colors = "viridis") %>% 
  layout(title = "Comparing Top 5 and Bottom 5 Human Freedom Scoring Countries",
      xaxis = list(title = "Human Freedom Score"),
      yaxis = list(title = "Happiness Score"))

```

This plot compares the top 5 and bottom 5 countries for human freedom scores. You can see that there is a good amount of variety in happiness scores among the top 5 and bottom 5 countries. Libya and Hong Kong have very similar happiness scores, which seems to indicate that high freedom scores do not necessarily mean that the population living in those countries are happier. Still, there is quite a divide in happiness scores across the other countries in the top 5 and bottom 5.

### Happiness Scores

#### Bottom 5 Countries for Happiness Scores
```{r, echo = F}
bot5_whr_merge %>% knitr::kable(digits = 3)
```

```{r}
bot5_whr %>% 
select(-happiness_score) %>% 
  pivot_longer(
    human_freedom_score:economic_freedom_score,
    names_to = "freedom_subgroup",
    values_to = "score"
  ) %>% 
  mutate(freedom_subgroup = fct_relevel(freedom_subgroup, "human_freedom_score", "personal_freedom_score", 
                     "economic_freedom_score")) %>% 
  ggplot(aes(x = countries, y = score, fill = freedom_subgroup)) + 
  geom_bar(stat = "identity", color = "black", position = position_dodge()) +
  ylim(min = 0, max = 10) + 
  ggtitle("Comparing Bottom 5 Countries on Freedom Subtypes") +
  xlab("Countries")+
  ylab("Freedom Score")+
  labs(fill = "Freedom Subgroup")+
  theme(axis.text.x = element_text(size = 8, angle = 25, hjust = 1))
```


Insert comments about bot 5 whr

#### Top 5 Countries for Happiness Scores
```{r, echo = F}
top5_whr_merge %>% knitr::kable(digits = 3)
```

```{r}
top5_whr %>% 
select(-happiness_score) %>% 
  pivot_longer(
    human_freedom_score:economic_freedom_score,
    names_to = "freedom_subgroup",
    values_to = "score"
  ) %>% 
  mutate(freedom_subgroup = fct_relevel(freedom_subgroup, "human_freedom_score", "personal_freedom_score", 
                     "economic_freedom_score")) %>% 
  ggplot(aes(x = countries, y = score, fill = freedom_subgroup)) + 
  geom_bar(stat = "identity", color = "black", position = position_dodge()) +
  ylim(min = 0, max = 10) + 
  ggtitle("Comparing Bottom 5 Countries on Freedom Subtypes") +
  xlab("Countries")+
  ylab("Freedom Score")+
  labs(fill = "Freedom Subgroup")
```


Insert comments about top 5 whr

```{r, message = F, warning=F}
whr_combo = full_join(bot5_whr, top5_whr)

whr_combo %>% 
  mutate(text_label = str_c(countries, "\nHappiness Score: ", happiness_score, "\nFreedom Score:", human_freedom_score)) %>%
  plot_ly(x = ~happiness_score, y = ~human_freedom_score, type = "scatter", mode = "markers", size = 3, alpha = 0.5, color = ~countries, text = ~text_label, colors = "viridis") %>% 
  layout( title = "Comparing Top 5 and Bottom 5 Happiness Scoring Countries",
      xaxis = list(title = "Happiness Score"),
      yaxis = list(title = "Human Freedom Score")
    )
```

Insert comments about whr plot
